---
title: "Bubble plot"
author: "Bela Hausmann (Joint Microbiome Facility)"
output:
  html_document:
    toc: yes
    toc_float:
      collapsed: no
      smooth_scroll: no
editor_options: 
  chunk_output_type: console
---

# remove all the objects from the R session
```{r}
rm(list=ls())#read input file, containing culture data
```

```{r}
source("0_common.R", chdir = TRUE)
library(mia)
library(RColorBrewer)
library(gridExtra)
library(tidyverse)
```


# Load container

```{r}
SE <- readRDS(file.path(RD_DIR, "SE_abs_all_finalremovedSamples.rds"))
#SE <- transformCounts(SE, method = "relabundance")
#SE_rel <- counts(SE) / rowSums(counts(SE))
SE_clr <- assay(SE, "clr")
SE_rel <- assay(SE, "relabundance")


SEtest <- SE[, SE$collection_date %in% c("0","5","11","20","35")]
#SEtest <- SE[, SE$collection_date %in% c("35")]

#SEtest <- SE[, SE$Species %in% c("UV7")]

```

#process data
```{r}
# Remove rows with Family values "NA" or "uncultured"
SEtest <- subset(SEtest, !(Family == "NA" | Family == "uncultured" | Family == "Unknown Family"))
# Get the 'Family' column from the rowData of 'SE'
family_column <- rowData(SEtest)$Family
domain_column <- rowData(SEtest)$Domain
# Extract unique names listed under 'Family'
unique_family_names <- unique(family_column)

# Print the unique family names
print(unique_family_names)
```

```{r}
# make custom order of the Samples
SEtest$Species <- factor(SEtest$Species, levels =c("SC","Ac", "E25","UV7"))

```


```{r}
#SEtest %>% colData() %>% as_tibble() %>% View()
```

# Prepare data

```{r}
Data <-
  SEtest %>%
  meltAssay(assay_name = "relabundance", add_row_data = TRUE, add_col_data = TRUE) %>%
  tidyr::unite("Lineage", Domain:ASV_ID, sep = " > ", remove = FALSE) %>%
  as.data.table()

#Data <- Data[relabundance > 0] # zeros are optional in bubble plots
Data <- Data[, relabundance := relabundance * 100] # convert to %
Data[, Lineage := Lineage %>% str_replace("( (NA|uncultured) >)+", " … >")]

# Select taxa for plot

selected_taxa <- getTopTaxa(SEtest, top = 100, method = "median", assay_name = "relabundance")
#selected_taxa <- getTopTaxa(SEtest, top = 999, method = "median", assay_name = "counts")
```


# Prepare data at family level
```{r}
# Aggregate ASVs by Family
library(dplyr)
# Aggregate ASVs by Family and sum up their counts
family_counts <- Data %>% 
  group_by(Family) %>% 
  summarize(Total_Count = sum(relabundance))
  #summarize(Total_Count = sum(counts))
# Get the top 10 families based on Total_Count
top_families <- family_counts %>%
  top_n(100, Total_Count) %>%
  pull(Family)

```

# Plot taxa

```{r}
plot_data <- Data[FeatureID %in% selected_taxa]
plot_data$Lineage %>%
  unique() %>%
  sort() %>%
  cat(sep = "\n")
# Filter plot_data to include only the top 10 families
plot_data_top10 <- plot_data[plot_data$Family %in% top_families, ]
```

#simple stacked Bar plot

```{r}
my_color = c("Acidobacteriaceae (Subgroup 1)" ="steelblue3", Actinospicaceae="pink", Alcaligenaceae ="red2", Catenulisporaceae = "orange2",Caulobacteraceae = "firebrick3",Devosiaceae = "#8dd3c7", Inquilinaceae="#aafae7", Micropepsaceae="chartreuse3", Micrcoccaceae="chartreuse", Microbacteriaceae="aquamarine4", Rhodanobacteraceae="chocolate3", Sphingobacteriaceae="#139696", Sphingomonadaceae = "darkviolet", Xanthobacteraceae = "#fb8072", Oxalobacteraceae="hotpink", Micrococcaceae="mediumslateblue", Streptomycetaceae="yellow2", Beijerinckiaceae="#90709a", Gemmatimonadaceae="#dbdfa2",Reyranellaceae = "coral", Acetobacteraceae = "#f5daa3",Paenibacillaceae = "#ccebc5" ,Micromonosporaceae = "brown1", Pseudomonadaceae = "red3", Chthoniobacteraceae= "chartreuse3", Solirubrobacteraceae= "#f2f2f2", Mycobacteriaceae= "brown", Rhizobiaceae = "lightblue1", Burkholderiaceae = "aquamarine", Chitinophagaceae = "#685582", Nocardioidaceae = "#fff397", Nocardiaceae = "steelblue", Pedosphaeraceae= "orange4", Vampirovibrionaceae = "chocolate4", Schlesneriaceae ="thistle", Dongiaceae = "#f2f", Enterobacteriaceae="yellow4", Parachlamydiaceae = "blue")


ggplot(Data, aes(x = User_sample_ID, y = relabundance, fill = Family)) +
  geom_bar(stat = "identity") +
  #scale_fill_manual(values= c("steelblue3", "orange2","firebrick3","thistle","chartreuse3","aquamarine4","chocolate3","cyan4","brown1","darkviolet","goldenrod1","hotpink","mediumslateblue","yellow","brown2"))+
  scale_fill_manual(values= my_color)+
  #scale_fill_brewer(palette= "Paired") +
  #theme(axis.text.x = element_text(angle = 45, hjust =1)) +
  theme_light() +
  theme(axis.text.x  = element_text(angle=0, vjust=0.5, size=10)) +
  theme(axis.title.x  = element_blank())+
  #theme(axis.text.y  = element_blank())+
  #theme(legend.position="none")+
  xlab("") +
  ylab("16S rDNA copies/g sediment") +
  #facet_grid(. ~ collection_date, scales = "free_y", space = "free_y", switch = "y") +
  facet_wrap(Species~., scales="free_y",nrow=8) +
   coord_flip() +
  xlab("") +
  ylab("abs abundance") 


ggsave(file.path(PLOTS_DIR, g("{PROJECT_NAME}_abs_barplot_T5.svg")), width = 6, height = 8, limitsize = FALSE)

```

# add column strategy to data for r or k strategy based on literature

```{r}
strategy = c("Acidobacteriaceae (Subgroup 1)" ="k", Caulobacteraceae = "r", Devosiaceae = "k", Microbacteriaceae="r", Rhodanobacteraceae="k", Sphingobacteriaceae="r",Sphingomonadaceae = "r",Xanthobacteraceae = "k", Streptomycetaceae="k", Beijerinckiaceae="k", Reyranellaceae = "k", Acetobacteraceae = "r", Rhizobiaceae = "r", Burkholderiaceae = "r", Chitinophagaceae = "r", Pseudomonadaceae ="r", Nocardioidaceae ="k")

# Convert strategy named vector to a data frame
strategy_df <- data.frame(
  Family = names(strategy),
  strategy = unname(strategy),
  stringsAsFactors = FALSE
)

# Join the strategy letter to plot_data_top10
plot_data <- Data %>%
  left_join(strategy_df, by = "Family")


Replicate_counts <- plot_data %>% 
  group_by(SampleID) %>% 
  summarize(Total_Count = sum(relabundance))
#Data<- Data %>%
 # left_join(strategy_df, by = "Family")
```


# plot strategies
```{r}

ggplot(plot_data, aes(x = User_sample_ID, y = relabundance, fill = strategy)) +
  geom_bar(stat = "identity") +
  #scale_fill_manual(values= c("steelblue3", "orange2","firebrick3","thistle","chartreuse3","aquamarine4","chocolate3","cyan4","brown1","darkviolet","goldenrod1","hotpink","mediumslateblue","yellow","brown2"))+
  #scale_fill_manual(values= my_color)+
  #scale_fill_brewer(palette= "Paired") +
  #theme(axis.text.x = element_text(angle = 45, hjust =1)) +
  theme_light() +
  theme(axis.text.x  = element_text(angle=0, vjust=0.5, size=10)) +
  theme(axis.title.x  = element_blank())+
  theme(axis.text.y  = element_blank())+
  #theme(legend.position="none")+
  xlab("") +
  ylab("16S rDNA copies/g sediment") +
  #facet_grid(. ~ collection_date, scales = "free_y", space = "free_y", switch = "y") +
  facet_wrap(Species~., scales="free_y",nrow=8) +
   coord_flip() +
  xlab("") +
  ylab("abs abundance") 


ggsave(file.path(PLOTS_DIR, g("{PROJECT_NAME}_strategy_barplot_T35.png")), width = 6, height = 8, limitsize = FALSE)

```


#########################################################################################################################
############################ Statistical analysis in a loop #############################################################
#########################################################################################################################

# Create new data.table with count
```{r}
strategy_table <- plot_data[, .(collection_date, Replicate, Species, strategy, relabundance)]
# Assuming your data is called dt
#condensed_counts <- strategy_table[, .(total_count = sum(counts, na.rm = TRUE)), 
 #                      by = .(collection_date, Species, strategy, Replicate)]
condensed_counts <- strategy_table[, .(total_count = relabundance), 
                       by = .(collection_date, Species, strategy, Replicate)]

counts_summary <- condensed_counts %>% # the names of the new data frame and the data frame to be summarised
    group_by(Species, collection_date, strategy) %>%   # the grouping variable
    summarise(mean_value = mean(total_count, na.rm = TRUE),  # calculates the mean of each group
              sd_value = sd(total_count, na.rm = TRUE), # calculates the standard deviation of each group
              n_value = sum(!is.na(total_count)),  # calculates the sample size per group
              SE_value = sd(total_count, na.rm = TRUE)/sqrt(n())) # calculates the standard error of each group
```

###### test for normality ######
```{r}
r_condensed_counts <- condensed_counts[condensed_counts$strategy == "r", c(1,2,3,4,5)]
k_condensed_counts <- condensed_counts[condensed_counts$strategy == "k", c(1,2,3,4,5)]

SC <- r_condensed_counts[r_condensed_counts$Species == "SC", c(1,2,3,4,5)]
inputSC <-  SC[SC$collection_date == "35",c(1,2,3,4,5)]
shapiro.test(inputSC$total_count) #  normal

AC <- r_condensed_counts[r_condensed_counts$Species == "Ac", c(1,2,3,4,5)]
inputAc <-  AC[AC$collection_date == "5",c(1,2,3,4,5)]
shapiro.test(inputAc$total_count) #  normal

E25 <- r_condensed_counts[r_condensed_counts$Species == "E25", c(1,2,3,4,5)]
inputE <-  E25[E25$collection_date == "5",c(1,2,3,4,5)]
shapiro.test(inputE$total_count) #   normal

UV7 <- r_condensed_counts[r_condensed_counts$Species == "UV7", c(1,2,3,4,5)]
inputUV7 <-  UV7[UV7$collection_date == "20",c(1,2,3,4,5)]
shapiro.test(inputUV7$total_count) #  normal


ggplot(inputAc, aes(x=total_count)) + 
  geom_density()
```

#################### Sort the data by strategy  #################################
```{r}
r_condensed_counts <- condensed_counts[condensed_counts$strategy == "r", c(1,2,3,4,5)]

library(dplyr)
library(rstatix)
library(tidyr)
library(purrr)

# Function to process one timepoint with Kruskal-Wallis + pairwise Wilcoxon
process_timepoint_nonparam <- function(r_condensed_counts, collection_date) {
  # Kruskal-Wallis test
  kruskal_res <- kruskal.test(total_count ~ Species, data = r_condensed_counts)
  kruskal_p <- kruskal_res$p.value
  
  # Pairwise Wilcoxon test (with BH correction)
  pwilcox <- pairwise.wilcox.test(r_condensed_counts$total_count, r_condensed_counts$Species, p.adjust.method = "BH")
  
  # Convert matrix of results into tidy dataframe
  pw_df <- as.data.frame(as.table(pwilcox$p.value)) %>%
    filter(!is.na(Freq)) %>%
    rename(Sample1 = Var1,
           Sample2 = Var2,
           Wilcox_pvalue = Freq) %>%
    mutate(
      Timepoint = collection_date,
      Kruskal_pvalue = kruskal_p,
      Wilcox_sig = case_when(
        Wilcox_pvalue < 0.001 ~ "***",
        Wilcox_pvalue < 0.01 ~ "**",
        Wilcox_pvalue < 0.05 ~ "*",
        TRUE ~ "ns"
      )
    ) %>%
    select(Timepoint, Sample1, Sample2, Kruskal_pvalue, Wilcox_pvalue, Wilcox_sig)
  
  return(pw_df)
}

# Apply to each timepoint
results_samplediff_nonparam <- r_condensed_counts %>%
  group_by(collection_date) %>%
  group_split() %>%
  map_df(~process_timepoint_nonparam(.x, unique(.x$collection_date)))

write.csv(results_samplediff_nonparam, "R_Stats_wilcox_results_sample_differences_all.csv", row.names = FALSE)
#############################################################################################################################

# Function to process one timepoint with Kruskal-Wallis + pairwise Wilcoxon
process_sample_nonparam <- function(r_condensed_counts, Species) {
  # Kruskal-Wallis test
  kruskal_res <- kruskal.test(total_count ~ collection_date, data = r_condensed_counts)
  kruskal_p <- kruskal_res$p.value
  
  # Pairwise Wilcoxon test (with BH correction)
  pwilcox <- pairwise.wilcox.test(r_condensed_counts$total_count, r_condensed_counts$collection_date, p.adjust.method = "BH")
  
  # Convert matrix of results into tidy dataframe
  pw_df <- as.data.frame(as.table(pwilcox$p.value)) %>%
    filter(!is.na(Freq)) %>%
    rename(Time1 = Var1,
           Time2 = Var2,
           Wilcox_pvalue = Freq) %>%
    mutate(
      Sample = Species,
      Kruskal_pvalue = kruskal_p,
      Wilcox_sig = case_when(
        Wilcox_pvalue < 0.001 ~ "***",
        Wilcox_pvalue < 0.01 ~ "**",
        Wilcox_pvalue < 0.05 ~ "*",
        TRUE ~ "ns"
      )
    ) %>%
    select(Sample, Time1, Time2, Kruskal_pvalue, Wilcox_pvalue, Wilcox_sig)
  
  return(pw_df)
}

# Apply to each timepoint
results_time_nonparam <- r_condensed_counts %>%
  group_by(Species) %>%
  group_split() %>%
  map_df(~process_sample_nonparam(.x, unique(.x$Species)))



write.csv(results_time_nonparam, "R_wilcox_Stats_results_change_over_time.csv", row.names = FALSE)
#############################################################################################################################

```

```{r}
k_condensed_counts <- condensed_counts[condensed_counts$strategy == "k", c(1,2,3,4,5)]

library(dplyr)
library(rstatix)
library(tidyr)
library(purrr)

# Function to process one timepoint with Kruskal-Wallis + pairwise Wilcoxon
process_timepoint_nonparam <- function(k_condensed_counts, collection_date) {
  # Kruskal-Wallis test
  kruskal_res <- kruskal.test(total_count ~ Species, data = k_condensed_counts)
  kruskal_p <- kruskal_res$p.value
  
  # Pairwise Wilcoxon test (with BH correction)
  pwilcox <- pairwise.wilcox.test(k_condensed_counts$total_count, k_condensed_counts$Species, p.adjust.method = "BH")
  
  # Convert matrix of results into tidy dataframe
  pw_df <- as.data.frame(as.table(pwilcox$p.value)) %>%
    filter(!is.na(Freq)) %>%
    rename(Sample1 = Var1,
           Sample2 = Var2,
           Wilcox_pvalue = Freq) %>%
    mutate(
      Timepoint = collection_date,
      Kruskal_pvalue = kruskal_p,
      Wilcox_sig = case_when(
        Wilcox_pvalue < 0.001 ~ "***",
        Wilcox_pvalue < 0.01 ~ "**",
        Wilcox_pvalue < 0.05 ~ "*",
        TRUE ~ "ns"
      )
    ) %>%
    select(Timepoint, Sample1, Sample2, Kruskal_pvalue, Wilcox_pvalue, Wilcox_sig)
  
  return(pw_df)
}

# Apply to each timepoint
results_samplediff_nonparam <- k_condensed_counts %>%
  group_by(collection_date) %>%
  group_split() %>%
  map_df(~process_timepoint_nonparam(.x, unique(.x$collection_date)))

write.csv(results_samplediff_nonparam, "K_Stats_wilcox_results_sample_differences_all.csv", row.names = FALSE)
#############################################################################################################################
k_condensed_counts$collection_date <- factor(k_condensed_counts$collection_date, levels =c("0","5","11","20","35"))

# Function to process one timepoint with Kruskal-Wallis + pairwise Wilcoxon
process_sample_nonparam <- function(k_condensed_counts, Species) {
  # Kruskal-Wallis test
  kruskal_res <- kruskal.test(total_count ~ collection_date, data = k_condensed_counts)
  kruskal_p <- kruskal_res$p.value
  
  # Pairwise Wilcoxon test (with BH correction)
  pwilcox <- pairwise.wilcox.test(k_condensed_counts$total_count, k_condensed_counts$collection_date, p.adjust.method = "BH")
  
  # Convert matrix of results into tidy dataframe
  pw_df <- as.data.frame(as.table(pwilcox$p.value)) %>%
    filter(!is.na(Freq)) %>%
    rename(Time1 = Var1,
           Time2 = Var2,
           Wilcox_pvalue = Freq) %>%
    mutate(
      Sample = Species,
      Kruskal_pvalue = kruskal_p,
      Wilcox_sig = case_when(
        Wilcox_pvalue < 0.001 ~ "***",
        Wilcox_pvalue < 0.01 ~ "**",
        Wilcox_pvalue < 0.05 ~ "*",
        TRUE ~ "ns"
      )
    ) %>%
    select(Sample, Time1, Time2, Kruskal_pvalue, Wilcox_pvalue, Wilcox_sig)
  
  return(pw_df)
}

# Apply to each timepoint
results_time_nonparam <- k_condensed_counts %>%
  group_by(Species) %>%
  group_split() %>%
  map_df(~process_sample_nonparam(.x, unique(.x$Species)))



write.csv(results_time_nonparam, "K_wilcox_Stats_results_change_over_time.csv", row.names = FALSE)
#############################################################################################################################

```

##################################
# parametric (ANOVA + tukey hsd)
##################################

```{r}
# Function to process one timepoint
process_timepoint <- function(k_condensed_counts, collection_date) {
  # Run ANOVA
  stat <- aov(total_count ~ Species, data = k_condensed_counts)
  anova_p <- summary(stat)[[1]][["Pr(>F)"]][1]
  
  # Tukey HSD with rstatix
  tukey <- tukey_hsd(stat)
  
  # Add timepoint & ANOVA p-value
  tukey %>%
    select(group1, group2, p.adj, p.adj.signif) %>%
    mutate(Timepoint = collection_date,
           Anova_pvalue = anova_p) %>%
    rename(Sample1 = group1,
           Sample2 = group2,
           TukeyHSD_pvalue = p.adj,
           TukeyHSD_sig = p.adj.signif) %>%
    select(Timepoint, Sample1, Sample2, Anova_pvalue, TukeyHSD_pvalue, TukeyHSD_sig)
}

# Apply for each timepoint
results_samplediff <- k_condensed_counts %>%
  group_by(collection_date) %>%
  group_split() %>%
  map_df(~process_timepoint(.x, unique(.x$collection_date)))

write.csv(results_samplediff, "K_Stats_results_sample_differences_all.csv", row.names = FALSE)
#############################################################################################################################

k_condensed_counts$collection_date <- factor(k_condensed_counts$collection_date, levels =c("0","5","11","20","35"))

process_sample <- function(k_condensed_counts, Species) {
  # ANOVA across timepoints
  stat <- aov(total_count ~ collection_date, data = k_condensed_counts)
  anova_p <- summary(stat)[[1]][["Pr(>F)"]][1]
  
  # Tukey HSD
  tukey <- tukey_hsd(stat)
  
  # Build results
  tukey %>%
    select(group1, group2, p.adj, p.adj.signif) %>%
    mutate(Sample = Species,
           Anova_pvalue = anova_p) %>%
    rename(Time1 = group1,
           Time2 = group2,
           TukeyHSD_pvalue = p.adj,
           TukeyHSD_sig = p.adj.signif) %>%
    select(Sample, Time1, Time2, Anova_pvalue, TukeyHSD_pvalue, TukeyHSD_sig)
}

results_time <- k_condensed_counts %>%
  group_by(Species) %>%
  group_split() %>%
  map_df(~process_sample(.x, unique(.x$Species)))

write.csv(results_time, "K_Stats_results_change_over_time.csv", row.names = FALSE)
#############################################################################################################################
```

```{r}
# Function to process one timepoint
process_timepoint <- function(r_condensed_counts, collection_date) {
  # Run ANOVA
  stat <- aov(total_count ~ Species, data = r_condensed_counts)
  anova_p <- summary(stat)[[1]][["Pr(>F)"]][1]
  
  # Tukey HSD with rstatix
  tukey <- tukey_hsd(stat)
  
  # Add timepoint & ANOVA p-value
  tukey %>%
    select(group1, group2, p.adj, p.adj.signif) %>%
    mutate(Timepoint = collection_date,
           Anova_pvalue = anova_p) %>%
    rename(Sample1 = group1,
           Sample2 = group2,
           TukeyHSD_pvalue = p.adj,
           TukeyHSD_sig = p.adj.signif) %>%
    select(Timepoint, Sample1, Sample2, Anova_pvalue, TukeyHSD_pvalue, TukeyHSD_sig)
}

# Apply for each timepoint
results_samplediff <- r_condensed_counts %>%
  group_by(collection_date) %>%
  group_split() %>%
  map_df(~process_timepoint(.x, unique(.x$collection_date)))

write.csv(results_samplediff, "R_Stats_results_sample_differences_all.csv", row.names = FALSE)
#############################################################################################################################

r_condensed_counts$collection_date <- factor(r_condensed_counts$collection_date, levels =c("0","5","11","20","35"))

process_sample <- function(r_condensed_counts, Species) {
  # ANOVA across timepoints
  stat <- aov(total_count ~ collection_date, data = r_condensed_counts)
  anova_p <- summary(stat)[[1]][["Pr(>F)"]][1]
  
  # Tukey HSD
  tukey <- tukey_hsd(stat)
  
  # Build results
  tukey %>%
    select(group1, group2, p.adj, p.adj.signif) %>%
    mutate(Sample = Species,
           Anova_pvalue = anova_p) %>%
    rename(Time1 = group1,
           Time2 = group2,
           TukeyHSD_pvalue = p.adj,
           TukeyHSD_sig = p.adj.signif) %>%
    select(Sample, Time1, Time2, Anova_pvalue, TukeyHSD_pvalue, TukeyHSD_sig)
}

results_time <- r_condensed_counts %>%
  group_by(Species) %>%
  group_split() %>%
  map_df(~process_sample(.x, unique(.x$Species)))

write.csv(results_time, "R_Stats_results_change_over_time.csv", row.names = FALSE)
#############################################################################################################################
```


```{r}
# Define timepoints and sample pairs
strategies <- c("r", "k")
# Define the timepoint subsets
timepoint_subsets <- list(c("0", "5"), c("0", "11"), c("0", "20"), c("0", "35"), c("5", "11"), c("5", "20"), c("5", "35"), c("11", "20"), c("11", "35"), c("20", "35"))


# Results table
results_strat <- data.table(
  strategy = character(),
  Species1 = character(),
  Species2 = character(),
  Anova_pvalue = numeric(),
  TukeyHSD_pvalue = numeric())


# Loop through timepoints and sample pairs
for (i in strategies) {
  for (subset in timepoint_subsets) {
    
    # Subset data for this timepoint and species pair
    subset_data <- condensed_counts[strategy == i & collection_date %in% subset]
  
    # Check that the subset contains both samples
    if (length(unique(subset_data$collection_date)) != 2) next
    # Perform the ANOVA
    subset_data$collection_date <- as.factor(subset_data$collection_date) # transform collection_date numbers to factors
    anova_result <- aov(total_count ~ collection_date, data = subset_data)
    # Perform Tukey's HSD test
    tukeyhsd_result <- TukeyHSD(anova_result)
    # Extract the p-value from the ANOVA and TukeyHSD results
    anova_pvalue <- summary(anova_result)[[1]][["Pr(>F)"]][1]
    tukeyhsd_pvalue <- tukeyhsd_result[[1]][, "p adj"][1]

    # Add a row to the results data.table with the summary statistics and significance level
    if (tukeyhsd_pvalue < 0.0001) {
      sig <- "****"
    } else if (tukeyhsd_pvalue < 0.001) {
      sig <- "***"
    } else if (tukeyhsd_pvalue < 0.01) {
      sig <- "**"
    } else if (tukeyhsd_pvalue < 0.05) {
      sig <- "*"
    } else {
      sig <- "ns"
    }
    # Add a row to the results data.table with the summary statistics
    results_strat <- rbind(results_strat, data.table(strategy = i, Species1 = subset[1], Species2 = subset[2], Anova_pvalue = anova_pvalue, TukeyHSD_pvalue = tukeyhsd_pvalue, TukeyHSD_sig = sig), fill = TRUE)
  }
}

# Save results as a CSV file
write.csv(results_strat, "Strat_TimeStats_UV7.csv", row.names = FALSE)
```


#################### Sort the data by Timepoint #################################
```{r}
# Define timepoints and sample pairs
strategies <- c("r", "k")
sample_subsets <- list(
  c("SC", "Ac"),
  c("SC", "E25"),
  c("SC", "UV7"),
  c("Ac", "E25"),
  c("Ac", "UV7"),
  c("E25", "UV7")
)

# Results table
results_strat <- data.table(
  strategy = character(),
  Species1 = character(),
  Species2 = character(),
  Anova_pvalue = numeric(),
  TukeyHSD_pvalue = numeric())


# Loop through timepoints and sample pairs
for (i in strategies) {
  for (subset in sample_subsets) {
    
    # Subset data for this timepoint and species pair
    subset_data <- condensed_counts[strategy == i & Species %in% subset]
  
    # Check that the subset contains both samples
    if (length(unique(subset_data$Species)) != 2) next
    # Perform the ANOVA
    anova_result <- aov(total_count ~ Species, data = subset_data)
    # Perform Tukey's HSD test
    tukeyhsd_result <- TukeyHSD(anova_result)
    # Extract the p-value from the ANOVA and TukeyHSD results
    anova_pvalue <- summary(anova_result)[[1]][["Pr(>F)"]][1]
    tukeyhsd_pvalue <- tukeyhsd_result[[1]][, "p adj"][1]

    # Add a row to the results data.table with the summary statistics and significance level
    if (tukeyhsd_pvalue < 0.0001) {
      sig <- "****"
    } else if (tukeyhsd_pvalue < 0.001) {
      sig <- "***"
    } else if (tukeyhsd_pvalue < 0.01) {
      sig <- "**"
    } else if (tukeyhsd_pvalue < 0.05) {
      sig <- "*"
    } else {
      sig <- "ns"
    }
    # Add a row to the results data.table with the summary statistics
    results_strat <- rbind(results_strat, data.table(strategy = i, Species1 = subset[1], Species2 = subset[2], Anova_pvalue = anova_pvalue, TukeyHSD_pvalue = tukeyhsd_pvalue, TukeyHSD_sig = sig), fill = TRUE)
  }
}

# Save results as a CSV file
write.csv(results_strat, "Strat_samplediff_Stats_T5.csv", row.names = FALSE)
```



